% le caratteristiche richieste dall'università sono elencate qui: https://stem.elearning.unipd.it/mod/book/view.php?id=234&chapterid=46#modalita
% 12pt: font richiesto dall'università
% twoside: i margini interni ed esterni sono scambiati per le pagine "a sinistra" e "a destra"
% openright: i capitoli cominciano in pagine dispari ("a destra")
% extreport: supporta 12pt
\documentclass[12pt,a4paper,twoside,openright]{extreport}

\usepackage{amsmath}                            % per avere più controllo sulle equazioni 
\usepackage{csquotes}                           % per le citazioni
\usepackage{enumitem}                           % per avere più controllo sulle enumerazioni
\usepackage[
    a4paper,
    top=2cm,bottom=2cm,
    outer=2cm,inner=3cm,
    includeheadfoot
]{geometry}                                     % margini richiesti dall'università
\usepackage{graphicx}                           % per le immagini
\usepackage{icomma}                             % per separare le cifre decimali con una virgola
\usepackage{minted}                             % per il codice con la colorazione della sintassi
\usepackage[a-1a]{pdfx}                         % formato richiesto dall'università
\usepackage[output-decimal-marker={,}]{siunitx} % per le unità di misura
\usepackage{subcaption}                         % per le sottodidascalie
\usepackage{svg}                                % per inserire file SVG

\usepackage[english]{babel}
\selectlanguage{english}

\usepackage[backend=bibtex,style=ieee]{biblatex}
\bibliography{bibliografia}

\usepackage{fontspec}
\setmainfont{Times New Roman} % carattere richiesto dall'università

\usepackage{setspace}
\onehalfspacing % interlinea richiesta dall'università

\sloppy % per evitare che il testo in \verb finisca oltre i margini

% questi valori vengono usati nella composizione del frontespizio
\title{Scalable Cloud Agnostic Architecture for IoT Data Analysis and Machine Learning}
\author{Discalzi Alessandro}
\date{05/12/2024}
\newcommand{\supervisor}{Prof. Vangelista Lorenzo}


\begin{document}
    \pagenumbering{roman}
    \pagestyle{empty} % per le prime pagine, non mostrare il numero di pagina

    \input{Capitoli/frontespizio}
    \cleardoublepage
    
    \input{Capitoli/dedica}
    \cleardoublepage

    \pagestyle{plain} % comincia a mostrare il numero di pagina

    % NOTA: l'ambiente \abstract rimuove il numero della pagina e resetta il contatore delle pagine. 
    \chapter*{Summary}
    The rapid expansion of the Internet of Things (IoT) has introduced significant challenges in managing and analyzing the vast amounts of data generated by interconnected devices. This thesis focuses on developing a scalable, cloud-agnostic architecture to efficiently handle data from IoT devices.

The core objective is to create an architecture that seamlessly integrates with various cloud platforms—such as Aruba Cloud, AWS, and Microsoft Azure—while supporting both horizontal and vertical scaling. The system must be capable of ingesting data via the MQTT protocol, managing data from online devices and archive data sources, and performing complex data preprocessing and analysis, including machine learning tasks. Security is a critical consideration, with the architecture designed to protect data both in transit and at rest.

To achieve these goals, the thesis involves a detailed examination of existing technologies, an analysis of different cloud providers, and the definition of a flexible base architecture. Practical testing on a real case study is used to validate the architecture’s performance and scalability. The proposed solution addresses the growing demands of IoT data management and provides a robust framework for future applications in diverse cloud environments.

    \cleardoublepage

    \tableofcontents
    \cleardoublepage
    
    \listoffigures
    \listoftables
    \cleardoublepage
    
    \pagenumbering{arabic}

    \input{Capitoli/introduction}
    \cleardoublepage
    \input{Capitoli/existing}
    \cleardoublepage
    \input{Capitoli/cloud}
    \cleardoublepage
    \input{Capitoli/method}
    \cleardoublepage
    \input{Capitoli/implementation}
    \cleardoublepage
    \input{Capitoli/tests}
    \cleardoublepage
    \input{Capitoli/conclusion}
    \cleardoublepage
    \input{Capitoli/appendixA}
    \cleardoublepage

    \printbibliography[heading=bibintoc]
\end{document}
